R- Assigment
========================================================

Exercise 1
-------------------------------------------------------
### 1.1
There is a really exciting data set included with the default R installation, but I can't remember what
it's called. I think it has something to do with chicken. Find it by keyword search and load it into your
workspace.

```{r eval=FALSE, echo=FALSE}
help.search("chicken")
data("chickwts")
```

### 1.2
Add a new column to chickwts that gives the weight in kilograms, rather than grams. Make sure the
column has a descriptive name.

```{r eval=FALSE, echo=FALSE}
chickwts$kg <- chickwts$weight/1000
```

### 1.3
Make a box-and-whiskers plot showing the distribution of chicken weight according to feed type. Make
sure to label the axes appropriately. (hint: use data$w ~ data$f form)

```{r eval=FALSE, echo=FALSE}
boxplot(chickwts$weight ~ chickwts$feed, xlab = "feed", ylab = "weight (g)")
```

Exercise 2
-------------------------------------------------------

### 2.1
Create a numeric vector x of length 50 that ranges from -pi to pi. Create a numeric vector y1 that is the
sine of x (in radians). Create a vector y2 that is the cosine of x.

```{r eval=FALSE, echo=FALSE}
x <- seq(-pi, pi, length = 50)
y1 <- sin(x)
y2 <- cos(x)
```

### 2.2
Plot y1 vs. x as a series of points joined by lines. On the same graph, add red-colored points for y2 vs.
x. Add a legend.

```{r eval=FALSE, echo=FALSE}
plot(x, y1, type = "b")
points(x, y2, col = "red")
legend("topleft", legend = c("sin(x)", "cos(x)"), col = c("black",
+ "red"), pch = 1, lty = c(1, NA), inset = 0.02)
```
### 2.3
Plot a histogram of y2, making sure there are enough bins to clearly see the trend. Plot a density curve
of y2 using the default parameters. Which plot is more faithful to the true distibution?

```{r eval=FALSE, echo=FALSE}
hist(y2, breaks = 20)
```

Exercise 3
---------------------------------------------

### 3.1
Download the [OVC Clinical Data](https://discovery.genome.duke.edu/express/resources/193/OVCclinicalinfo.xls). Now, import the spreadsheet as a data frame into your R
workspace, naming the resulting object "clin". Briefy inspect the data.

```{r eval=FALSE, echo=FALSE}
library(gdata)
clin <- read.xls("OVCclinicalinfo.xls")
Converting xls file to csv file... Done.
Reading csv file... Done.
str(clin)
```

### 3.2
Confirm that the last three columns are useless, and remove them. Convert the first column to character
type. Change the name of the third column to "event". Convert the "CA125.POST" and "GRADE"
columns into numeric values, with the ambiguous entries coded as NAs. Fix the "Debulk" column such
that only two levels are used ("O" and "S"). Rename the eighth column to "response" and convert it to
a logical vector (0 = FALSE, 1 = TRUE).

```{r eval=FALSE, echo=FALSE}
clin <- clin[, 1:8]
clin$OVC.TumorID <- as.character(clin$OVC.TumorID)
colnames(clin)[3] <- "event"
clin$CA125.POST <- as.numeric(as.character(clin$CA125.POST))
clin$GRADE <- as.numeric(as.character(clin$GRADE))
clin$Debulk <- factor(substr(clin$Debulk, 1, 1))
colnames(clin)[8] <- "response"
clin$response <- as.logical(clin$response)
str(clin)
```

### 3.3
Now that you have "cleaned up" the "clin" object, save it for later use, both as an R object ("clin.rda")
and also as a CSV file ("clin.csv").

```{r eval=FALSE, echo=FALSE}
save(clin, file = "clin.rda")
write.csv(clin, file = "clin.csv")
```

