\documentclass{article}
\begin{document}


\title{DNA sequence assembly project}

\maketitle

Assignment: for an Illumina ecoli dataset, a Roche 454 ecoli dataset and a mixed 454/Illumina Toxoplasma gondii dataset, clean the sequencing reads, assemble the genomes, map the reads back to genomes, find variants (if any), determine closest relative (by sequence similarity), annotate (computationally).\\
Data can be found on the central server:
\begin{verbatim}
/mnt/home/uirig/user_data/BCB504
\end{verbatim}

Submit one latex/sweave report for the entire project (ie. part 2 should continue part 1, etc.). The report should include full details of the analysis and the results, as if written in a publication. Finally, include a section that details individual contribution to the analysis/report. 


\section*{Part 1: Clean the reads}
\textbf{Due Wed April 2nd}\\
\begin{itemize}
\item Perform Quality Assurance of the reads
\item Removing, or trim, poor quality reads
\item Use of R/Python, sfffile, sffinfo
\end{itemize}

\section*{Part 2: Assemble the genomes}
\textbf{Due Fri April 13th}\\
\begin{itemize}
\item Assembly and optimize the assembly (maximize the metrics)
\item What are the details about the assembly (genome size, coverage, illumina insert sizes, etc.)
\end{itemize}

\section*{Part 3: Map the reads, find variants, annotate}
\textbf{Due Fri April 22nd}\\
\begin{itemize}
\item Map the reads back to the assembly.
\item report the details (\% of reads that map).
\item are there variants.
\item What is the closest public genome.
\item annotate the genomes. 
\end{itemize}


\end{document}