\documentclass{article}
\usepackage{Sweave}
\begin{document}



\section*{Exercise 1}
\begin{itemize}
\item Generate a random DNAString of length 2000. (You will need: sample(), DNA\_BASES, paste(), DNAString().)
\item Create views on it.
\item Invert the views.
\item Count the frequencies of the DNA letters: (a) in the DNAString object, (b) inside the views, (c) outside the views. Do a sanity check.
\end{itemize}

\section*{Exercise 2}
\begin{itemize}
\item Load Affymetrix hgu95av2 probe sequences into a DNAStringSet object.
\item Remove the first 10 probes.
\item Which probes contain more than 16 A's?
\item Reverse complement the probes.
\item Trim the first (5') and last (3') two bases.
\item Generate the sequences of the mismatch probes (MM probes) by replacing the middle nucleotide of each PM probe by its reverse complement.
\item Which probes contain more than 9 consecutive A's? (You can use vcount- Pattern() for this.) Display their sequences.
\end{itemize}

\section*{Exercise 3}
\begin{itemize}
\item Load BSgenome data package for hg19.
\item Count the number of times each Affymetrix hgu95av2 probe hits Human chr22. (You will need: PDict(), countPDict().)
\item Which probes have more than 2000 hits? Display their sequences.
???\end{itemize}

\section*{Exercise 4}
The goal of this exercise is to count the nb of times each Human transcript is hit by a hgu95av2 probe. We use the TxDb.Hsapiens.UCSC.hg19.knownGene package for the locations of the transcripts and their exons.
\begin{itemize}
\item Extract the Human transcriptome with extractTranscriptsFromGenome() (defined in the GenomicFeatures package).
\item Use vcountPDict() (with 'collapse=2') to count the nb of hits per tran-script.
\end{itemize}


\end{document}
